---
title: "Creating a new document using pandocfilters"
author: "Florian Schwendinger and Andrie de Vries"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Creating a new document using pandocfilters}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


The constructor functions in the **pandocfilters** package can also be used to create a new document by reproducing the AST. To show how this is done we use the first paragraph of chapter 2.1 from the [R-FAQ](https://cran.r-project.org/doc/FAQ/R-FAQ.txt)  and apply different inline elements. The output is saved as HTML file. 

In general, a **pandocfilters** document is a list of two elements:     

* Metadata
* Content
   
The content is composed of lists of blocks, where each block itself usually  consists of a list of inline elements. 

**NOTE:**

*If an inline object is needed and a character vector is given, **pandocfilters** automatically transforms the character string to an 'inline' object of type 'Str'.*
    

## Creating a new document

```{r initialize}
require("pandocfilters", quietly = TRUE, warn.conflicts = FALSE)
```



```{r}
# Create a new document.
doc <- document()

# Create a non-standard writer function so we can look at the document while writing it.
cat_writer <- function(x, con = stdout(), format = "json") {
  x <- pandoc_from_json(x, to = format)
  cat(paste(x, collapse = "\n"))
}

# Append a Header and look at the document
args(doc$append_header)
```


```{r}
doc$append_header( "R Basics" )
doc$write(format = "html", writer = cat_writer)
```


```{r, results='hide'}
# Append a level 2 Header
doc$append_header( "What is R?", level = 2)

# Append Plain text with inline formating
x <- c(
  Emph("R"), Space(), "is a system for ", 
  Strong("statistical computation"), Space(), Strikeout("and"), 
  Space(), Superscript("graphics"), ". ", LineBreak(), 
  Subscript("It"), Space(), SmallCaps("consists"), Space(), 
  Quoted("of", quote_type = "SingleQuote"), Space(), 
  Quoted("a", quote_type = "DoubleQuote"), Space(), 
  RawInline("html", "<i>language</i>"), 
  " plus a run-time environment with", " graphics, a debugger, access to ",
  "certain system functions,", 
  " and the ability to run programs stored in script files."
)

doc$append_plain( x )
doc$write(format = "html", writer = cat_writer)
```


<div style="background-color:powderblue; margin:10%;">
```{r, results='asis', echo = FALSE}
doc$write(format = "html", writer = cat_writer)
```
</div>

## Modifying the abstract syntax tree

If you carefully at the output, you can see that quotes don't look very nice. Instead, they look like garbage characters.

This can be fixed by using the function `astrapply()`. In most cases it is best to look at the AST via `str()` to see what should be replaced. Since the quote doesn't look good in the HTML case we replace it with `<q>text</q>`.


```{r, results='hide'}
fix_quotes_fun <- function(x) {
  RawInline("html", sprintf("<q>%s</q>", x$c))
}

fix_quotes <- function(type, content, ...) {
  if (type == "Quoted") {       
    lapply(content[[-1]], fix_quotes_fun)
  }
}
```


Now modify the document and inspect the results:

```{r, results='hide'}
doc$doc <- astrapply(doc$doc, FUN = fix_quotes)
doc$write(format = "html", writer = cat_writer)
```

<div style="background-color:powderblue; margin:10%;">
```{r, results='asis', echo=FALSE}
doc$write(format = "html", writer = cat_writer)
```
</div>

## Create a Table

The *pandocfilters* package makes it easy to construct tables in pandoc format.

```{r}
table <- document()
table$append_table(cars[1:3, ])
```

Once you have the table, you can render the output to various different formats.

### markdown


```{r}
table$write(format = "markdown", writer = cat_writer)
```

### html

```{r}
table$write(format = "html", writer = cat_writer)
```

### html5

```{r}
table$write(format = "html5", writer = cat_writer)
```

### org

```{r}
table$write(format = "org", writer = cat_writer)
```

### latex

```{r}
table$write(format = "latex", writer = cat_writer)
```


### rst

```{r}
table$write(format = "rst", writer = cat_writer)
```


### asciidoc
```{r}
table$write(format = "asciidoc", writer = cat_writer)
```

