---
title: "Introduction to pandocfilters"
author: "Florian Schwendinger and Andrie de Vries"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to pandocfilters}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


The document converter [pandoc](http://pandoc.org/) is widely used
in the R community. One feature of pandoc is that it can produce and consume
JSON-formatted abstract syntax trees (AST). This allows to transform a given
source document into JSON-formatted AST, alter it by so called filters and pass the altered JSON-formatted AST back to pandoc. This package provides functions which allow to write such filters in native R code. The package is inspired by the Python package [pandocfilters](https://github.com/jgm/pandocfilters/).

To alter the AST, the JSON representations of the data structures building the AST have to be replicated. For this purpose, **pandocfilters** provides a set of constructors, with the goal to ease building / altering the AST.

## Installation 

You can find information about installing pandoc at 
[http://pandoc.org/installing.html](http://pandoc.org/installing.html).  

**NOTE:** 
*There have been API-changes in pandoc 1.16. To account for this changes, 
**pandocfilters** needs to know the pandoc version. Therefore, **pandocfilters**   will try to get the pandoc version at startup. However, it is also possible 
to get and set the pandoc version via the functions `get_pandoc_version()` 
and `set_pandoc_version()`.*


```{r initialize}
require("pandocfilters", quietly = TRUE, warn.conflicts = FALSE)
# Get the pandoc version.
get_pandoc_version()
```


```{r pandoc_version, eval=FALSE}
# Set the pandoc version.
set_pandoc_version(1.16)
```


## Constructors

As mentioned before, constructors are used to replicate the pandoc AST in R.
For this purpose, pandoc provides two basic types, **inline** elements and 
**block** elements. An extensive list can be found below.

To minimize the amount of unnecessary typing **pandocfilters** automatically 
converts character strings to pandoc objects of type `"Str"` if needed.
Furthermore, if a single inline object is provided where a list of inline 
objects is needed **pandocfilters** automatically converts this inline 
object into a list of inline objects.

For example, the canonical way to emphasize the character string `"some text"` would be

```r
Emph(list(Str("some text")))
```

Since single inline objects are automatically transformed to lists of inline objects, this is equivalent to

```r
Emph(Str("some text"))
```

Since a character string is automatically transformed to an inline object, this is equivalent to

```r
Emph("some text")
```

In short, whenever a list of inline objects is needed one can also use a single inline object or a character string, and therefore the following three code lines are equivalent.   


```{r emph, eval = FALSE}
Emph(list(Str("some text")))
Emph(Str("some text"))
Emph("some text")
```


### Inline Elements

1. `Str(x)`
2. `Emph(x)`
3. `Strong(x)`
4. `Strikeout(x)`
5. `Superscript(x)`
6. `Subscript(x)`
7. `SmallCaps(x)`
8. `Quoted(x, quote_type)`
9. `Cite(citation, x)`
10. `Code(code, name, language, line_numbers, start_from)`
11. `Space()`
12. `SoftBreak()`
13. `LineBreak()`
14. `Math(x)`
15. `RawInline(format, x)`
16. `Link(target, text, title, attr)`
17. `Image(target, text, caption, attr)`
18. `Span(attr, inline)`
               
        
### Block Elements

1. `Plain(x)`
2. `Para(x)`
3. `CodeBlock(attr, code)`
4. `BlockQuote(blocks)`
5. `OrderedList(lattr, lblocks)`
6. `BulletList(lblocks)`
7. `DefinitionList(x)`
8. `Header(x, level, attr)`
9. `HorizontalRule()`
10. `Table(rows, col_names, aligns, col_width, caption)`
11. `Div(blocks, attr)`
12. `Null()`

### Argument Constructors

1. `Attr(identifier, classes, key_val_pairs)`
2. `Citation(suffix, id, note_num, mode, prefix, hash)`
3. `TableCell(x)`


## Altering the AST

To read, write and test the AST you can use the some utility functions: 

### Utility Functions


```{r utility_functions, eval=FALSE}
pandoc_to_json()

pandoc_from_json()

test_filter()
```

## Examples

### Lower Case 

The following example shows how to obtain the AST from a markdown file 
(`"lower_case.md"`) and convert every object of type `Str` to lower case.


```{r caps}
caps <- function(key, value, ...) {
    if (key == "Str") return( Str( tolower(value) ) )
    return(NULL)
}

example_1 <- file.path(
  system.file(package = "pandocfilters"), 
  "examples", 
  "lower_case.md"
)

# the file before transformation
readLines(example_1)
```


```{r apply_caps}
# read connection
input_connection <- textConnection(pandoc_to_json(example_1, from="markdown"))
# write connection
output_connection <- textConnection("modified_ast", open="w")

# apply filter
filter(caps, input=input_connection, output=output_connection)

# convert altered ast to markdown
pandoc_from_json(modified_ast, to="markdown")
```


```{r close}
close(input_connection)
close(output_connection)
```

